---
title: "qdapRegex R Notebook"
output:
  html_document:
    df_print: paged
---

Folks, 

Recall the regex from session 1? We'd figured out ways to match regular pattewrns in text expressions.

Over time, people have collaborated on systematizing and standardizing some of the regex codes developed. 

Below I present R package `qdapRegex` that draws on a few such (mainly following US conventions). 

## first, setup chunk

```{r}
rm(list=ls())  # clear workspace
 if (!require(qdapRegex)) {install.packages("qdapRegex")}

 library(qdapRegex)
 require(magrittr)

```

### Extract Twitter Hash Tags, Name Tags, & URLs

We can write regex code to capture twitter hastags, name tags and URLs. However, `ex_hash()` and sister functions in *qdapRegex* help ease the process quite a bit. Behold.

```{r}
 x <- c("@hadley I like #rstats for #ggplot2 work.",
    "Difference between #magrittr and #pipeR, both implement pipeline operators for #rstats:
        http://renkun.me/r/2014/07/26/difference-between-magrittr-and-pipeR.html @timelyportfolio",
    "Slides from great talk: @ramnath_vaidya: Interactive slides from Interactive Visualization
        presentation #user2014. http://ramnathv.github.io/user2014-rcharts/#1"
  )
 x  # view raw text
 
 ex_hash(x)  # extract all hashtags
 ex_tag(x)  # extract name tags
 ex_url(x)  # extract URLs

```

The Q had come up earlier that whereas `\d` or `[0-9]` matches numbers, what about decimals? fractions? Etc?

```{r}
 # Extract Numbers
 z <- c("-2 is an integer.  -4.3 and 3.33 are not.",
    "123,456 is a lot more than -.2",
    "hello world -.q")
 z
 rm_number(z)  # remove all no.s
 ex_number(z)  # extract no.s from text
 ex_number(z) %>% as_numeric()

```

### Extract Time entities

Another common match use-case for regex is catching different time formats. Here too, a list of common patterns are coded into a nice wrapper func `ex_time()`:
```{r}
 # Extract Times
 x <- c(
    "I'm getting 3:04 AM just fine, but...",
    "for 10:47 AM I'm getting 0:47 AM instead.",
    "no time here",
    "Some time has 12:04 with no AM/PM after it",
    "Some time has 12:04 a.m. or the form 1:22 pm"
    )
 x
 ex_time(x)
 ex_time(x) %>% as_time()
 ex_time(x) %>% as_time(., as.POSIXlt = TRUE)

```

### Phone numbers

And yup, where else to go but match phone number formats. Mostly US style convention - maybe its time we build one for Indian number formats too?

```{r}
 # Extract phone numbers
 x <- c(" Mr. Bean bought 2 tickets 2-613-213-4567 or 5555555555 call either one",
  "43 Butter Rd, Brossard QC K0A 3P0 - 613 213 4567", 
  "Please contact Mr. Bean (613)2134567",
  "1.575.555.5555 is his #1 number",  
  "7164347566",
  "I like 1234567 dogs"
  )
 x # view example text

 rm_phone(x)  # remove all phone no.s
 ex_phone(x)  # extract all phone no.s

```

### Extract bracketed text

Any text inside brackets of whatever variety can be matched and extracted. Easy to do this via regular regex too I guess but still, here goes...

```{r}
 # Extract Bracketed Text
 y <- c("I love chicken [unintelligible]!", 
    "Me too! (laughter) It's so good.[interrupting]",
    "Yep it's awesome {reading}.", "Agreed. {is so much fun}")
 
  y  # view example text

  ex_bracket(y)   

```

Another cool use-case is extracting citations from academic or research texts.

```{r}
 # Extract Citations
 w <- c("Hello World (V. Raptor, 1986) bye (Foo, 2012, pp. 1-2)",
    "Narcissism is not dead (Rinker, 2014)",
    "The R Core Team (2014) has many members.",
    paste("Bunn (2005) said, \"As for elegance, R is refined, tasteful, and",
        "beautiful. When I grow up, I want to marry R.\""),
    "It is wrong to blame ANY tool for our own shortcomings (Baer, 2005).",
    "Wickham's (in press) Tidy Data should be out soon.",
    "Rinker's (n.d.) dissertation not so much.",
    "I always consult xkcd comics for guidance (Foo, 2012; Bar, 2014).",
    "Uwe Ligges (2007) says, \"RAM is cheap and thinking hurts\"",
    "Silly (Bar, 2014) stuff is what Bar (2014, 2012) said."
  )
 w   # view example text
 
 ex_citation(w) 
 ex_citation(w) %>% as_count()

```

Well, that's it for now. 

Most of the examples are admittedly taken directly from Tyler Rinker's Git page for the qdapRegex package.

Sudhir